Storing messages with multipartite neural cliques

### 2014-07-23 Articles/article.tex
...

	We study the properties of the network once some messages have been stored in it. We are interested in how much edges have been added to the network and if the process introduced confusion between messages. We also compute the rate of the network in the sense of information theory (ratio between information stored in the network and the information used to store the network)

@@

	Theoretical and empirical densities match very closely, see Figure \ref{densiteth}. %gap of theoretical density with density in simulations : under $1$ percent, 
	The higher $a$ is, the quicker density approaches $1$.

	Multipartite cliques saturate quicker, which was expected since they take much more edges.

@@

	The rate $R$ of the network measures how much information is stored in the network compared to what it takes to store the network itself. Note it doesn't take into account how effective messages retrieving is (if we store too much messages in the network, the proportion of retrieved messages approaches $0$ but the rate is high)

@@

	\subsection{Accepting random messages}

@@

	We are interested in distinct problems of messages retrieval. We detail what those problems are. Since those problems vary in hardness, we propose distinct measures for efficiencies. Efficiency measures the ratio of information effectively retrieved times the hardness of the problem. It can be expressed as the product of the rate of the network $R$ (how much information is stored), the retrieval rate $P_{retrieve}$ (how much information is effectively retrieved) and some measure of hardness of the problem $h$ (how hard the problem really is).
	\begin{align}
		\eta = P_{retrieve} R \cdot h
		\label{defeff}
	\end{align}

	To account for hardness of problems, we consider an "optimal" associative memory. In this document, we are interested in retrieving messages with a few iterations at most. However an "optimal" memory could store $\mathcal{M}$ with $H(\mathcal{M})$ bits and recover messages by maximum likelihood (with ambiguities causing errors in our study, to simplify the problem). For example, such a memory would scan over all messages to find all messages corresponding to a partial probe. We detail the performance of such a memory for each problem, which gives use a measure for the hardness of each problem.

@@

	We compute the probability for a message $m$ and a partial probe $\tilde{m}$ to engender confusion.

@@

	%simulations faites avec erreurs non symétriques !!

@@

	\begin{align}
		h_{resilience} = h_{erasures} \times \left( 1 - H(\psi) \right)^{-1}
	\end{align}	
	, where $H(\psi) = -(\psi \log_2 (\psi) + (1-\psi) \log_2 (1-\psi)$

@@

	Thanks to the theoretical formula for density (\ref{formula_density}), we compute analytical error rates after one iteration of the retrieving rule. We compare those results to empirical error rates obtained through simulations\footnote{Simulations are done by taking means on pool of several networks and messages sets, and each point is drawn for different networks and messages sets.}. We also compare efficiencies of distinct networks for distinct values of the parameter $a$ (weight of the constant weight code) in the sense of information theory and show improvements over the case where $a = 1$.

@@

	One can see a partial version of a message $\mu$ as the result of erasures of symbols of $x$. Some symbols of $x$ are replaced by the blank character $\intercal$.	Erasures of a symbol in a message corresponds to an "erased" cluster in the network. An erased cluster is a cluster of units which are all inactive, that is to mean set to $0$.

@@

	Whence error rate in retrieving messages is : 
	\begin{align}	
		P_{err} = 1 - \left(1 - d^	{a(c-c_e)}\right)^{c_e(\ell-a)} 
	\end{align}

	In order to compare distinct networks using distinct alphabets (due to variations of the parameter $a$), we define the efficiency of the network relatively to the hardness of the problem (how much it is efficient compared to a perfect associative memory). A perfect associative memory would use exactly the same number of bits than the original messages and retrieve messages by maximum likelihood (ambiguities causing errors). Efficiency is the amount of information retrieved divided by the hardness of the problem.

	The network is a binary symmetric matrix with null diagonal and can therefore be stored in memory as $\frac{c(c-1) \ell^2}{2}$ bits. Since messages are not stored in an ordered fashion, the entropy of the set of messages is $\log_2(\frac{({\ell \choose a}^c)^M}{M!}) \mathop{=}_{M \rightarrow +\infty} M(c \log_2{\ell \choose a } - \log_2(M) + \frac{1}{\log 2} + o(1)) $ (thanks to  Stirling's formula). The ratio of information effectively stored in the network is $P_{retrieve} \times  \frac{2M \left(c \log_2{\ell \choose a } - \log_2(M) + \frac{1}{\log 2} \right)}{c(c-1)\ell^2}$.

	Ambiguities would occur in a perfect associative memory if at least two messages were identical over non-erased letters. The probability it doesn't happen is $ (1-{\ell \choose a}^{c - c_e})^{M-1}$.

	$\eta_M = P_{retrieve}  \frac{2 M\left(c \log_2{\ell \choose a } - \log_2(M) + \frac{1}{\log 2} \right)}{c(c-1)\ell^2} \times (1-(\frac{1}{{\ell \choose a}})^{c - c_e})^{-(M-1)}$
	%\frac{c}{(c-c_e) \log_2{l \choose a }}

	The maximum ratio of information that can be stored in the network is therefore $\eta = \max_x \eta_x $, which we will define as the efficiency of the network.

@@

		Simulations agree with the analytical results for error rate, albeit we can observe a small shift, see Figure \ref{erasuresth}. The theoretical error rate is a bit optimistic which is due to the connections not being independent.

		 For one iteration, as long as the parameter $w$ is between $1$ and $a$, there is no effect on retrieving (in this setting selecting amounts to take the maximum as all units of the original message achieve the same score). However for $4$ iterations, the choice of $\alpha = a$ allows a shift of about $5000$ messages stored on top of those stored for $\alpha = 1$. The variant rule can thus be considered a significant improvement over the classic one for multipartite clique networks (recall that $\alpha = 1$ amounts to the classic retrieving rule).

@@

	Messages can suffer many types of alteration. Multipartite clique neural networks can also correct errors.

@@

	%Correctly activated units in uncorrupted clusters achieve at least a score of $a(c-c_e - 1) + \gamma$. 
	Being connected to a given activated unit in a corrupted cluster occurs with probability $d$, since errors are identically distributed. Moreover corrects units see their scores shift, as do all activated units respectively thanks to being part of the original message and the memory effect $\gamma$. We can express those probability laws as : 
	For example an \textbf{a}ctivated \textbf{g}ood unit in a \textbf{c}orrect cluster starts with a base score of $ a(c - c_e - 1) + \gamma$. The probability it has to be connected to a random activated unit in an erroneous cluster is $d$. Thereby the law for its score $s_{agc}$ is :

@@

	$P_{err} = 1 - P_{retrieve}$

	\subsubsection{Simulations}
	See Figure \ref{corruptth} compares the two choices $\alpha = 1$ and $\alpha = a$.
	"$a$-winners take all" better than "winner takes all" with errors instead of erasures, even for one iteration.

@@

		$\eta_m = P_{retrieve}  \frac{2 M\left(c \log_2{\ell \choose a } - \log_2(M) + \frac{1}{\log 2} \right)}{c(c-1)\ell^2} \times \left [\sum_{k = c_e+1}^{c} \left (1-\frac{1}{{\ell \choose a}} \right )^k \left (\frac{1}{{\ell \choose a}} \right )^{c-k} \right ] ^{-(M-1)}$

@@

	%gamma = infini meilleur sur une itération
	We assume $\gamma = \infty$ for one iteration as it prevents activated neurons to be deactivated.
	With probability $1 - \psi$, one correct vertex is still connected to a given activated unit in an intact cluster. Thus $P(n_{v_c} = x ) = {a (c_k-1) \choose x} (1-\psi)^{x} \psi ^ { a (c - c_e -1) - x }$

@@

	The probability for an edge to be present in the network after damages is : $P_+ = \psi (1 - d) + (1 - \psi) d$. So for other units $P(n_v = x) = {a (c - c_e) \choose x} P_+^x (1-P_+)^{a (c - c_e) -x }$.

@@

	The probability for all correct units to achieve a score equal or greater than $x$, with at least one achieving a score of $x$ is $P_c (x) =\sum_{k = 1}^{a} { a \choose k }   P(n_{v_c} = x)^k P(n_{v_c} > x)^{a-k} $

	%If $\gamma$ is big enough,
	It follows that $P(\mbox{no other vertex activated in one cluster})= \sum_{x = 1}^{a (c - c_e)} P_c(x) P(n_v < x)^{\ell-a}$.%[ \sum_{x = 0}^{n - 1} P(n_v = x) \right]^{l-a }$

@@

	\begin{align}
	P_{retrieve}=  \left( \sum_{x = 1}^{a ( c- c_e)} P_c(x) P(n_v < x)^{\ell-a} \right )^{c_e}
	\label{psi_formula_th}
	\end{align}

@@

	With random noise on edges, hardness of the problem must take into account the capacity of the binary symmetric channel (the closer $\psi$ is of $0.5$, the harder the problem is). Our new is efficiency is written as :  	
	$\eta_{m, \psi} = P_{retrieve}  \frac{2 M \left(c \log_2{\ell \choose a } - \log_2(M) + \frac{1}{\log 2} \right)}{c(c-1)l^2} \times (1-{\ell \choose a}^{c - c_e})^{-(M-1)} \times \frac{1}{1 + \psi \log_2(\psi) +(1-\psi) \log_2(1-\psi)} $

	$\eta_\psi = max_x \, \eta_{x, \psi}$
	\begin{figure}[!htb]
		\includegraphics[width=8.5cm]{Courbes/fig2psic8l256e4psi2percent}%thpsi_c8l256e4}
		\caption{$\psi = 0.02$, 2 activities per cluster, 8 clusters, 256 units per cluster, 4 erasures, each point is the mean of 10 networks with 1000 sampled messages, analytical result in continuous black}
		\label{psith}
	\end{figure}

>>>>

	We study the properties of the network once some messages have been stored in it. We are interested in how much edges have been added to the network %and if the process introduced confusion between messages. 
	We also compute the rate of the network in the sense of information theory (ratio between information stored in the network and the information used to store the network)

@@

	Theoretical and empirical densities match very closely, see Figure \ref{densiteth}. %gap of theoretical density with density in simulations : under $1$ percent, 
	The higher $a$ is, the quicker density approaches $1$ : multipartite cliques saturate quicker, which was expected since they take much more edges ($a^2 \binom{c}{2}$).

@@

	The rate $R$ of the network measures how much information is stored in the network compared to what it takes to store the network itself. Note it doesn't take into account how effective retrieving messages is (if we store too much messages in the network, the proportion of retrieved messages approaches $0$ but the rate is high)

@@

	%\subsection{Accepting random messages}

@@

	\label{pstat}

	We are interested in distinct problems of messages retrieval. We detail what those problems are. Since those problems vary in hardness, we propose distinct measures for efficiencies. Efficiency measures the ratio of information effectively retrieved times the hardness of the problem. It can be expressed as the product of the rate of the network $R$ (how much information is stored, c.f. (\ref{rateth})), the retrieval rate $P_{retrieve}$ (how much is effectively retrieved) and some measure of hardness of the problem $h$ (how hard the problem is).
	\begin{align}
		\eta_M = P_{retrieve} R\, h
		\label{defeff}
	\end{align}
	However what we are really interested in is the maximum as we want to exploit the network at its best. The maximum efficiency attainable with a given retrieving rule, which we will call efficiency, is thereby \begin{align}
		\eta = \max_{M \ge 1} \eta_M
		\label{defmaxef}
	\end{align}
	To account for hardness of problems, we consider an "optimal" associative memory. In this document, we are interested in retrieving messages with a few iterations at most. However an "optimal" memory would store $\mathcal{M}$ with $H(\mathcal{M})$ bits and recover messages by maximum likelihood (with ambiguities causing errors in our study, to simplify the problem). For example, such a memory would scan over all messages to find all messages corresponding to a partial probe. We detail the performance of such a memory for each problem, which gives use a measure for the hardness of each problem.

@@

	We compute the probability for a message $m$ and a partial probe $\tilde{m}$ to cause confusion.

@@

	%simulations faites avec erreurs non symétriques !!
	% mais apparemment cela joue très peu (de toute façon l'erreur théorique concordait avec le résultat des simulations (je suis en train de les refaire)

@@

	\begin{align}
		h_{resilience} = h_{erasures} \times \left( 1 - H(\psi) \right)^{-1}
	\end{align}	
	, where $H(\psi) = -(\psi \log_2 (\psi) - (1-\psi) \log_2 (1-\psi)$.

@@

	Thanks to the theoretical formula for density (\ref{formula_density}), we compute analytical error rates after one iteration of the retrieving rule. We compare those results to empirical error rates obtained through simulations\footnote{Simulations are done by taking means on pool of several networks and messages sets, and each point is drawn for different networks and messages sets.}. We also compare efficiencies of distinct networks for distinct values of the parameter $a$ (weight of the constant weight code) in the sense of information theory and show improvements over the case where $a = 1$. Those efficiencies are computed using the empirical retrieval rate and formulas of section \ref{pstat}.

@@

	%One can see a partial version of a message $\mu$ as the result of erasures of symbols of $x$. Some symbols of $x$ are replaced by the blank character $\intercal$.	Erasures of a symbol in a message corresponds to an "erased" cluster in the network. An erased cluster is a cluster of units which are all inactive, that is to mean set to $0$.

@@

	Whence error rate in retrieving messages is : 
	\begin{align}	
		P_{err} = 1 - P_{retrieve} = 1 - \left(1 - d^	{a(c-c_e)}\right)^{c_e(\ell-a)} 
	\end{align}

	%In order to compare distinct networks using distinct alphabets (due to variations of the parameter $a$), we define the efficiency of the network relatively to the hardness of the problem (how much it is efficient compared to a perfect associative memory). A perfect associative memory would use exactly the same number of bits than the original messages and retrieve messages by maximum likelihood (ambiguities causing errors). Efficiency is the amount of information retrieved divided by the hardness of the problem.

	%The network is a binary symmetric matrix with null diagonal and can therefore be stored in memory as $\frac{c(c-1) \ell^2}{2}$ bits. Since messages are not stored in an ordered fashion, the entropy of the set of messages is $\log_2(\frac{({\ell \choose a}^c)^M}{M!}) \mathop{=}_{M \rightarrow +\infty} M(c \log_2{\ell \choose a } - \log_2(M) + \frac{1}{\log 2} + o(1)) $ (thanks to  Stirling's formula). The ratio of information effectively stored in the network is $P_{retrieve} \times  \frac{2M \left(c \log_2{\ell \choose a } - \log_2(M) + \frac{1}{\log 2} \right)}{c(c-1)\ell^2}$.

	%Ambiguities would occur in a perfect associative memory if at least two messages were identical over non-erased letters. The probability it doesn't happen is $ (1-{\ell \choose a}^{c - c_e})^{M-1}$.

	%In terms of efficiency, 
	%$\eta_M = P_{retrieve}  \frac{2 M\left(c \log_2{\ell \choose a } - \log_2(M) + \frac{1}{\log 2} \right)}{c(c-1)\ell^2} \times (1-(\frac{1}{{\ell \choose a}})^{c - c_e})^{-(M-1)}$
	%\frac{c}{(c-c_e) \log_2{l \choose a }}

	%The maximum ratio of information that can be stored in the network is therefore $\eta = \max_x \eta_x $, which we will define as the efficiency of the network.

@@

		Simulations agree with the analytical results for error rate, albeit we can observe a small shift, see Figure \ref{erasuresth}. The theoretical error rate is slightly optimistic which is due to the connections not being independent.

		 For one iteration, as long as the parameter $\alpha$ is between $1$ and $a$, there is no effect on retrieving (in this setting selecting amounts to take the maximum as all units of the original message achieve the same score). However for $4$ iterations, the choice of $\alpha = a$ allows a shift of almost $5000$ messages stored in addition to those stored for $\alpha = 1$. The variant rule can thus be considered as a significant improvement over the classic one for multipartite clique networks (recall that $\alpha = 1$ amounts to the classic retrieving rule).

@@

%	Messages can suffer many types of alteration. Multipartite clique neural networks can also correct errors.

	We now examine the case of retrieving messages from corrupt probes.

@@

	%Correctly activated units in uncorrupted clusters achieve at least a score of $a(c-c_e - 1) + \gamma$.
	We consider $a$ to be small in regard to $\ell$ which allows us to ignore the possibility for a random replacement cluster to share common activated units with the original.

	Corrects units see their scores shift, as do all activated units respectively thanks to being part of the original message and the memory effect $\gamma$.
	Being connected to a given activated unit in a corrupted cluster occurs with probability $d$, since errors are identically distributed. Moreover  We can express those probability laws as : 
	For example an \textbf{a}ctivated \textbf{g}ood unit in a \textbf{c}orrect cluster starts with a base score of $ a(c - c_e - 1) + \gamma$. The probability it has to be connected to a random activated unit in an erroneous cluster is $d$. Thereby the law for its score $s_{agc}$ is :

@@

	We can of course compute the error rate : $P_{err} = 1 - P_{retrieve}$

	\subsubsection{Simulations}
	Figure \ref{corruptth} compares the two choices $\alpha = 1$ and $\alpha = a$. The choice of $\alpha = a$ is a considerable improvement over $\alpha = 1$ for retrieving messages from corrupt probe.
	%"$a$-winners take all" better than "winner takes all" with errors instead of erasures, even for one iteration.

	The analytical error rate is close to the empirical one, albeit slightly optimistic.

@@

		%$\eta_m = P_{retrieve}  \frac{2 M\left(c \log_2{\ell \choose a } - \log_2(M) + \frac{1}{\log 2} \right)}{c(c-1)\ell^2} \times \left [\sum_{k = c_e+1}^{c} \left (1-\frac{1}{{\ell \choose a}} \right )^k \left (\frac{1}{{\ell \choose a}} \right )^{c-k} \right ] ^{-(M-1)}$

	We now study the influence of the parameter $a$ on retrieving messages from partial probes after the network has sustained damages.

@@

	%gamma = infini meilleur sur une itération
	We take $\gamma = \infty$ (or big enough) for one iteration as it prevents activated neurons to be deactivated. Another good choice is $\gamma = a$ in practice.

	With probability $1 - \psi$, a correct vertex is still connected to a given activated unit in an intact cluster. Thus $P(s_{g} = x ) = {a (c_k-1) \choose x} (1-\psi)^{x} \psi ^ { a (c - c_e -1) - x }$

@@

	The probability for an edge to be present in the network after damages is : $P_+ = \psi (1 - d) + (1 - \psi) d$. So for other units $P(s_b = x) = {a (c - c_e) \choose x} P_+^x (1-P_+)^{a (c - c_e) -x }$.

@@

	The probability for all correct units to achieve a score equal or greater than $x$, with at least one achieving a score of $x$ is $P_g (x) =\sum_{k = 1}^{a} { a \choose k }   P(s_g = x)^k P(s_g > x)^{a-k} $

	%If $\gamma$ is big enough,
	It follows that $P(\mbox{no other vertex activated in one cluster})= \sum_{x = 1}^{a (c - c_e)} P_g(x) P(s_b < x)^{\ell-a}$.%[ \sum_{x = 0}^{n - 1} P(n_v = x) \right]^{l-a }$

@@

	\begin{align}
	P_{retrieve}=  \left( \sum_{x = 1}^{a ( c- c_e)} P_g(x) P(s_b < x)^{\ell-a} \right )^{c_e}
	\label{psi_formula_th}
	\end{align}

@@

%	With random noise on edges, hardness of the problem must take into account the capacity of the binary symmetric channel (the closer $\psi$ is of $0.5$, the harder the problem is). Our new is efficiency is written as :  	
%	$\eta_{m, \psi} = P_{retrieve}  \frac{2 M \left(c \log_2{\ell \choose a } - \log_2(M) + \frac{1}{\log 2} \right)}{c(c-1)l^2} \times (1-{\ell \choose a}^{c - c_e})^{-(M-1)} \times \frac{1}{1 + \psi \log_2(\psi) +(1-\psi) \log_2(1-\psi)} $
%	
%	$\eta_\psi = max_x \, \eta_{x, \psi}$
	\begin{figure}[!htb]
		\includegraphics[width=8.5cm]{Courbes/fig2psic8l256e4psi2percent}%thpsi_c8l256e4}
		\caption{$\psi = 0.02$, 2 activities per cluster, 8 clusters, 256 units per cluster, 4 erasures, each point is the mean of 10 networks with 1000 sampled messages, analytical result in continuous black}
		\label{psith}
	\end{figure}
